Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-19 paper/results.tex
Updated counts of iterations.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@667 3fc257bb-8550-0410-81b3-a5c103da83f7

In Figure~\ref{f:cc}, we show the performance of the connected
component identification algorithm (Figure~\ref{fig:cc}).  A
comparison is made with a hybrid ``Giant Connected Component''
implementation using both Trilinos and MR-MPI.  Power law graphs often
have one or two very large components and many very small components.
The hybrid algorithm exploits this feature by using an inexpensive
breadth-first search (BFS) from the vertex with highest degree to
identify the largest components one at a time, followed by a more
expensive algorithm to identify the small components in the remainder
of the graph.  In our hybrid implementation, we intially perform
matrix-vector multiplications in Trilinos to perform a BFS, finding a
handful of components that include 70\% or more of the vertices (for
these R-MAT matrices).  We then apply our MapReduce algorithm to the
remaining graph to identify the small components.  This hybrid
approach is quite effective, reducing the execution time to identify
all components in RMAT-20 and RMAT-24 by 80-99\%, as shown in
Figure~\ref{f:cc}.  For RMAT-28, where the graph is too large to fit
into memory, Trilinos cannot perform the initial BFS, so only the
MapReduce algorithm was used.  The total number of components
identified were A, B, C for the 3 increasing R-MAT sizes; the number
of iterations required by the algorithm to find them was X, Y, Z
respectively.

@@

The execution times for the maximal independent set (MIS) algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the
R-MAT generation results, superlinear speed-up of the algorithm occurs
for RMAT-20 and RMAT-24, as more of the graph fits into processor
memory and less file I/O is needed.  For RMAT-28, the algorithm
requires significant out-of-core operations. In this case, parallel
efficiency is nearly perfect going from 8 to 64 processors.  The
number of vertices in the resulting MIS were A, B, C for the 3
increasing R-MAT sizes; the number of iterations required by the
algorithm to find them was X, Y, Z respectively.

@@

The execution times for the SSSP algorithms of Figures~\ref{fig:sssp}
and \ref{fig:sssp2} are shown in Figure~\ref{f:sssp}.  The results
show the benefit of using the enhanced algorithm, which gave at least
a 17\% (and often greater) reduction in execution time due to reduced
communication; only the updated distances are communicated throughout
most of the enhanced algorithm.  However, the execution times are
still large compared to multi-threaded implementations; for example,
Madduri et al.~\cite{Madduri07} report execution times of only 11
seconds on 40 processors for a multithreaded implementation on a Cray
MTA-2 (the predecessor of the XMT) on graphs with one billion edges.
The number of iterations required in the MapReduce algorithms to find
the SSSP to all vertices for the 3 increasing R-MAT sizes was X, Y, Z
respectively.

>>>>

In Figure~\ref{f:cc}, we show the performance of the connected
component identification algorithm (Figure~\ref{fig:cc}).  A
comparison is made with a hybrid ``Giant Connected Component''
implementation using both Trilinos and MR-MPI.  Power law graphs often
have one or two very large components and many very small components.
The hybrid algorithm exploits this feature by using an inexpensive
breadth-first search (BFS) from the vertex with highest degree to
identify the largest components one at a time, followed by a more
expensive algorithm to identify the small components in the remainder
of the graph.  In our hybrid implementation, we intially perform
matrix-vector multiplications in Trilinos to perform a BFS, finding a
handful of components that include 70\% or more of the vertices (for
these R-MAT matrices).  We then apply our MapReduce algorithm to the
remaining graph to identify the small components.  This hybrid
approach is quite effective, reducing the execution time to identify
all components in RMAT-20 and RMAT-24 by 80-99\%, as shown in
Figure~\ref{f:cc}.  For RMAT-28, where the graph is too large to fit
into memory, Trilinos cannot perform the initial BFS, so only the
MapReduce algorithm was used.  The total number of components
identified were roughly 498K, 9.37M, and 169M for the three increasing 
R-MAT sizes; five or six iterations were 
required by the algorithm.

@@

The execution times for the maximal independent set (MIS) algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the
R-MAT generation results, superlinear speed-up of the algorithm occurs
for RMAT-20 and RMAT-24, as more of the graph fits into processor
memory and less file I/O is needed.  For RMAT-28, the algorithm
requires significant out-of-core operations. In this case, parallel
efficiency is nearly perfect going from 8 to 64 processors.  The
number of vertices in the resulting MIS were 368K, 5.18M, and 72.4M
for the three
increasing R-MAT sizes; the number of iterations required by the
algorithm ranged from five for RMAT-20 to nine for RMAT-28.

@@

The execution times for the SSSP algorithms of Figures~\ref{fig:sssp}
and \ref{fig:sssp2} are shown in Figure~\ref{f:sssp}.  The results
show the benefit of using the enhanced algorithm, which gave at least
a 17\% (and often greater) reduction in execution time due to reduced
communication; only the updated distances are communicated throughout
most of the enhanced algorithm.  However, the execution times are
still large compared to multi-threaded implementations; for example,
Madduri et al.~\cite{Madduri07} report execution times of only 11
seconds on 40 processors for a multithreaded implementation on a Cray
MTA-2 (the predecessor of the XMT) on graphs with one billion edges.
Seven to ten iterations were required in the MapReduce algorithms to find
the SSSP to all vertices.
